Wigglyhedra

### 2024-05-28 draft/wigglyhedra.tex
Added reference and some more writing about KS complexes.

The triangulated category of interest will be the category \(\mathcal{C}_m = D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
The algebra \(A_m\) is defined as the quotient of the path algebra of the quiver with \(m+1\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    0 \arrow[bend left]{r}& 1 \arrow[bend left]{r} \arrow[bend left]{l}& 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The doubled quiver of the type \(A_5\) Dynkin graph.}
  \label{fig:am-quiver}
\end{figure}

Recall that the type \(A_m\) Dynkin graph is an undirected graph with \(m\) vertices and \(m-1\) edges; each edge connects successive pairs of points as shown in~\cref{fig:a5-dynkin}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet
  \end{tikzcd}
  \caption{The type \(A_5\) Dynkin graph.}
  \label{fig:a5-dynkin}
\end{figure}
Given an undirected graph as above, we can construct its \emph{doubled quiver}, which is a directed graph in which every undirected edge is replaced by a pair of oppositely oriented arrows, as shown in~\cref{fig:a5-doubled}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    \bullet \arrow[bend left]{r}& \bullet \arrow[bend left]{r} \arrow[bend left]{l}& \bullet \arrow[bend left]{r}\arrow[bend left]{l}& \bullet \arrow[bend left]{r}\arrow[bend left]{l}& \bullet \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The doubled quiver of the type \(A_5\) Dynkin graph.}
  \label{fig:a5-doubled}
\end{figure}
The \emph{zig-zag algebra} of an undirected graph is a finite-dimensional quotient of the path algebra of the doubled quiver of that graph (see~\cite[Section 3]{hue.kho:01} for the precise definition).

Let \(Z_m\) be the zig-zag algebra of the type \(A_m\) graph.
Let \(\mathcal{C}_m\) be \(K^b(Z-\text{proj})\), which is the bounded homotopy category of graded projective modules over \(Z_m\)
The category \(\mathcal{C}_m\) is triangulated.
The precise definition of this category appears in, e.g.,~\cite{sei.tho:01,tho:06} for type \(A_m\) and in~\cite{bap.deo.lic:20} for general undirected graphs.

The category \(\mathcal{C}_m\) is linear over a fixed field \(k\), triangulated, and 2-Calabi--Yau.
That is, for every pair of objects \(X\) and \(Y\) in \(\mathcal{C}_m\), we have
\[\Hom_{\mathcal{C}_m}(X,Y[d]) \cong \Hom_{\mathcal{C}_m}(Y, X[2-d]),\]
where \([1]\) denotes the triangulated shift.
\asilata{Update these definitions to the bigraded version, not the 2-CY version.}
Furthermore, \(\mathcal{C}_m\) is strongly generated by \(m\) objects denoted by \(P_1, \ldots, P_m\), which correspond to indecomposable projective modules over the zig-zag algebra.
These objects satisfy:
\begin{itemize}
\item \(\Hom(P_i,P_i[d]) = 0\) if \(d \neq 0, 2\);
\item \(\Hom(P_i,P_i[d]) = k\) if \(d = 0, 2\);
\item \(\Hom(P_i, P_{i\pm1}[d]) = 0\) if \(d \neq 1\);
\item \(\Hom(P_i, P_{i\pm1}[1]) = k\), and
\item \(\Hom(P_i, P_j[d]) = 0 \) otherwise.
\end{itemize}

Now consider \(n+2\) marked points on a disk, arranged on a straight horizontal line, and labelled \(0,1, \ldots, n+1\).
We will consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we also refer to such an isotopy class as a \emph{curve}.
The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

In particular, the line segment joining point \(i\) to point \(i+1\) is sent to the object \(P_{i+1}\) in the category \(\mathcal{C}_{n+1}\).
It is easy to check that the (minimal) intersection numbers between any two such segments counts the number of morphisms between the associated objects.
In particular, the self-intersection number of a single segment is computed to be \(2\), as the second copy of the segment can be slightly deformed to be disjoint from the first copy except at the two endpoints.
In fact, it is true for any pair of curves that its minimal intersection number is the number of morphisms between the associated objects (\cite[Lemma 4.11]{kho.sei:02}).
\asilata{Now figure out what to say about \(t\)-structures, namely that wiggly arcs are exactly those whose associated complexes lie in the heart of the standard \(t\)-structure.}

@@

>>>>

The category of interest is a subcategory of \(D^b(A_m\text{-mod})\) for a certain finite-dimensional algebra \(A_m\).
That is, a subcategory of the bounded derived category of (finitely generated) \(A_m\)-modules.
The algebra \(A_m\), called the \emph{zigzag algebra}, is defined as a quotient of the path algebra of the quiver \(Q_m\) with \(m\) vertices depicted in~\cref{fig:am-quiver}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    1 \arrow[bend left]{r} & 2 \arrow[bend left]{r}\arrow[bend left]{l}& \cdots \arrow[bend left]{r}\arrow[bend left]{l}& m \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The quiver \(Q_m\). The algebra \(A_m\) is a quotient of the path algebra \(kQ_m\).}
  \label{fig:am-quiver}
\end{figure}
Let \(kQ_m\) be the path algebra of \(Q_m\), graded by path length.
Let \(e_i\) denote the idempotent path at the vertex \(i\), and let \((i|j)\) denote the arrow from vertex \(i\) to a (neighbouring) vertex \(j\).
A longer path, which is a composition of arrows, will be denoted by \((i_1|i_2|\cdots|i_n)\).

The zigzag algebra \(A_m\) is the quotient of \(kQ_m\) by the following relations:
\begin{itemize}
\item set \((i|j|l) = 0\) whenever \(i \neq l\);
\item set \((i|j|i) = (i|l|i)\) for every \(j\) and \(l\) such that there are arrows between \(i\) and \(j\) and between \(i\) and \(l\).
\end{itemize}
See~\cite[Section 3]{hue.kho:01} for further details, and~\cite[Remark 6.6]{bap.deo.lic:20} for a more detailed discussion.
Let \(P_i\) denote the indecomposable projective \(A_m\)-module defined as \(P_i = A_m\cdot e_i\).
For any integer \(j\), let \(P_i \langle j \rangle\) denote its graded shift by \(j\).
Set \(\mathcal{C}_m\) to be the smallest triangulated subcategory of \(D^b(A_m\text{-mod})\) containing all the objects \(P_i\langle j \rangle\) (see, e.g.,~\cite[Section 2.3.3]{bap.deo.lic:22} for further discussion).

The category \(\mathcal{C}_m\) is \(k\)-linear and triangulated, where \([1]\) denotes the triangulated shift.
For \(X, Y \in \mathcal{C}_m\), let \(\Hom^{i,j}(X,Y)\) denote the space of morphisms in \(\mathcal{C}_m\) of homological degree \(i\) and internal degree \(j\).
In other words, we have
\[\Hom^{i,j}(X,Y) = \Hom(X,Y\langle i \rangle[j]).\]
Furthermore, for any integers \(a\) and \(b\), we have
\[\Hom^{i,j}(X\langle a \rangle[b], Y\langle a \rangle[b]) = \Hom^{i,j}(X,Y).\]

\begin{remark}\label{rem:generating-morphisms}
  The generating objects \(P_i\) (and hence their graded and homological shifts) satisfy the following:
  \begin{itemize}
  \item \(\Hom^{a,b}(P_i,P_i) = 0\) if \(b \neq 0\) and \(a \neq 0, 2\);
  \item \(\Hom^{a,0}(P_i,P_i) = k\) if \(a = 0, 2\);
  \item \(\Hom^{a,b}(P_i, P_{i\pm1}) = 0\) if \(b \neq 0\) and \(a \neq 1\);
  \item \(\Hom^{1,0}(P_i, P_{i\pm1}) = k\), and
  \item \(\Hom^{a,b}(P_i, P_j) = 0 \) otherwise.
  \end{itemize}
\end{remark}

The category \(\mathcal{C}_m\) has a standard bounded \(t\)-structure, whose heart consists of the \emph{linear complexes}.
\asilata{Say more about linear complexes.}

\subsection{The Khovanov--Seidel curve model}
Now consider \(n+2\) marked points on a disk, arranged on a straight horizontal line, and labelled \(0,1, \ldots, n+1\).
We consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we also refer to such an isotopy class as a \emph{curve}.
The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading\asilata{Should this say homological?} shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

In particular, the line segment joining point \(i\) to point \(i+1\) is sent to the object \(P_{i+1}\) in the category \(\mathcal{C}_{n+1}\).
It is easy to check that the (minimal) intersection numbers between any two such segments counts the number of morphisms\asilata{Discuss (or choose to ignore) grading.} between the associated objects.
In particular, the self-intersection number of a single segment is computed to be \(2\), as the second copy of the segment can be slightly deformed to be disjoint from the first copy except at the two endpoints.
In fact, it is true for any pair of curves that its minimal intersection number is the number of morphisms between the associated objects (\cite[Lemma 4.11]{kho.sei:02}).

\begin{lemma}
Every wiggly arc corresponds to an object in the heart of the standard \(t\)-structure.  
\end{lemma}
\begin{proof}
  The procedure to build up the complex of projective modules from a curve explicitly describes the differentials.
  Each component of these differentials is a map from a shift of \(P_i\) to a shift of \(P_j\).
  It can be seen from~\cref{rem:generating-morphisms} that any component that maps from \(P_i\) to \(P_j\) for \(i \neq j\) keeps the complex linear; it is only when we have a looping map from \(P_i\) to \(P_i\langle 2 \rangle\) that we move out of the standard heart.
  \asilata{Now figure out what to say about \(t\)-structures, namely that wiggly arcs are exactly those whose associated complexes lie in the heart of the standard \(t\)-structure.}
\end{proof}

@@

%%% Local variables:
%%% TeX-command-extra-options: "-pvc"
%%% End:
